### 2019-04-06 ReplyToReferee2019-Apr.tex
Updates from Overleaf

\documentclass[letter]{article}
\usepackage{bbm,mathtools,array,longtable,booktabs,graphicx,color,enumitem}

\textwidth 6.5in
\textheight 9in
\hoffset -1in
\voffset -1in
\input MACROS.tex
\begin{document}
\title{Reply to the Referees}
\maketitle

\subsection*{Report beginning ``The submitted paper \ldots''}

\begin{enumerate}
    \item[5.] The norm for $\calF$ has been identified.
\end{enumerate}

\subsection*{Report beginning ``This paper proposes \ldots''}

\begin{enumerate}[labelwidth = 10ex]
    \item[$6^{\text{th}}$ \bullet] The norm does depend on the $\lambda_i$ because the $u_i$ depend on the $\lambda_i$.  This is clearer not that the norm for $\calF$ has been identified.
\end{enumerate}

\subsection*{Report beginning ``This article studies \ldots''}

\begin{enumerate}
    \item[??] 
\end{enumerate}

\end{document}